The \dealii Library, Version 9.2

### 2020-05-28 9.2/paper.tex
add myself to 9.2
brief discreption of the DiscreteTime class is added

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
     Bruno Blais,
     Thomas C. Clevenger,
%     Denis Davydov,
    Marc Fehling,
%     Daniel Garcia-Sanchez,
    Alexander V. Grayver,
%     Graham Harper,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Peter Munch,
%     Ross Maguire Kynch,
    Matthias Maier,
%     Jean-Paul Pelteret,
    Bruno Turcksin,
%     David Wells
    Zhuoran Wang,
  },
  pdftitle={The deal.II Library, Version 9.2, 2020},
}

@@

%
% \author[4]{Jean-Paul~Pelteret}
%
\author[*1]{Bruno~Turcksin}
%
% \author[12]{David Wells}
% \affil[12]{Department of Mathematics, University of North Carolina,
%   Chapel Hill, NC 27516, USA.
%   {\texttt{drwells@email.unc.edu}}}

@@

The major changes of this release are:
%
\begin{itemize}
\item xy \todo[inline]{Update once we have the subsections in Section 2;
  provide cross-references to each of these subsections}
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}),
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item \dealii{} had decent support for solving complex-valued problems
  (e.g., ones in quantum mechanics -- like the example covered by the
  \texttt{step-58} tutorial program covered below -- or for
  time-harmonic problems) for a while already. However, there were two
  areas in which support was missing. First, the UMFPACK direct solver
  packaged with \dealii{} did not support solving complex-valued
  linear problems. This has now been addressed: UMFPACK actually can
  solve such systems, we just needed to write the appropriate
  interfaces. Second, the \texttt{DataOut} class that is responsible
  for converting nodal data into information that can then be written
  into files for visualization did not know how to deal with vector-
  and tensor-valued fields whose components are complex numbers. An
  example for this is to solve the time-harmonic version of the
  Maxwell equations that has the electric and magnetic fields as
  solution. This, too, has been addressed in this release.
  \item \todo[inline]{Reza: Please add a short summary of the
    \texttt{DiscreteTime} class}
\item z \todo[inline]{What else? Maybe mention the updated step-12?}
\end{itemize}
%
In addition to these changes, the changelog lists more than 200 other
features and bugfixes.

@@

By replacing the collective communications during set up and removing the arrays
that contain information for each process (enabled by the application of consensus
algorithms and other modifications---a full list of modifications leading to this
improvement can be found online), we were able to significantly
improve the set up
time for large-scale simulations and to solve a Poisson problem with multigrid
with \num{2.1e12} unknowns.
Figure~\ref{fig:init_costs} compares the timings of simulations of various problem
sizes (including set up) on 49,152 MPI ranks using a matrix-free solver according
to~\cite{KronbichlerKormann2019,KronbichlerWall2018} discontinuous elements of
degree $5$ in a geometric multigrid (GMG) scheme with the previous release 9.1
and with the current release 9.2. While the times for the V-cycle have shown
excellent scaling before, many initialization routines have been considerably
improved, especially the enumeration of unknowns on the multigrid levels and
the setup of the multigrid transfer.

@@

\begin{figure}
  \begin{tikzpicture}
    \begin{loglogaxis}[
      width=0.51\textwidth,
      height=0.38\textwidth,
          xlabel={\# DoFs},
          ylabel={time [s]},
          title={9.1 release},
          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
          tick label style={font=\scriptsize},
          xticklabel style={rotate=25,anchor=north east},
          label style={font=\scriptsize},
          legend style={font=\scriptsize},
          %legend to name=legend:scalingMG,
          %legend cell align=left,
          %legend columns = 3,
          ymin=2e-3,ymax=8e1,
          grid, thick]
          \addplot table[x index = 0, y index={1}] {\tableinitold};
          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinitold};
          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinitold};
          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinitold};
          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinitold};
          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinitold};
    \end{loglogaxis}
  \end{tikzpicture}
  \hfill
  \begin{tikzpicture}
    \begin{loglogaxis}[
      width=0.51\textwidth,
      height=0.38\textwidth,
          xlabel={\# DoFs},
          ylabel={time [s]},
          title={9.2 release},
          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
          tick label style={font=\scriptsize},
          xticklabel style={rotate=25,anchor=north east}, thick,
          label style={font=\scriptsize},
          legend style={font=\scriptsize},
          legend to name=legend:init,
          legend cell align=left,
          legend columns = 6,
          ymin=2e-3,ymax=8e1,
          grid]
          \addplot table[x index = 0, y index={1}] {\tableinit};
          \addlegendentry{create mesh};
          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinit};
          \addlegendentry{enumerate (mg) dofs};
          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinit};
          \addlegendentry{init matrix-free};
          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinit};
          \addlegendentry{init GMG transfer};
          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinit};
          \addlegendentry{init smoother};
          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinit};
          \addlegendentry{GMG V-cycle};
    \end{loglogaxis}
  \end{tikzpicture}
  \\
  \strut\hfill\pgfplotslegendfromname{legend:init}\hfill\strut
  \caption{Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right).}
  \label{fig:init_costs}
\end{figure}

@@

\todo[inline]{Reza: Short paragraph about constexpr}

\todo[inline]{Mention as last sentence that next release will use C++14.}

@@

R.~M.~Kynch was supported by the Engineering and Physical Science Research
Council (EPSRC) UK through grant EP/K023950/1 while working at Swansea
University, UK 2013-2015 where the majority of his contribution was
undertaken.

>>>>

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
     Bruno Blais,
     Thomas C. Clevenger,
%     Denis Davydov,
    Marc Fehling,
%     Daniel Garcia-Sanchez,
    Alexander V. Grayver,
%     Graham Harper,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Peter Munch,
%     Ross Maguire Kynch,
    Matthias Maier,
%     Jean-Paul Pelteret,
    Reza Rastak,
    Bruno Turcksin,
    David Wells
    Zhuoran Wang,
  },
  pdftitle={The deal.II Library, Version 9.2, 2020},
}

@@

%
% \author[4]{Jean-Paul~Pelteret}
%
 \author[14]{Reza Rastak}
 \affil[14]{Department of Civil and Environmental Engineering,
   Stanford University,
   Stanford, CA 94305, USA.
   {\texttt{rastak@stanford.edu}}}

\author[*1]{Bruno~Turcksin}
%
\author[12]{David Wells}
\affil[12]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

The major changes of this release are:
%
\begin{itemize}
\item xy \todo[inline]{Update once we have the subsections in Section 2;
  provide cross-references to each of these subsections}
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}),
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item \dealii{} had decent support for solving complex-valued problems
  (e.g., ones in quantum mechanics -- like the example covered by the
  \texttt{step-58} tutorial program covered below -- or for
  time-harmonic problems) for a while already. However, there were two
  areas in which support was missing. First, the UMFPACK direct solver
  packaged with \dealii{} did not support solving complex-valued
  linear problems. This has now been addressed: UMFPACK actually can
  solve such systems, we just needed to write the appropriate
  interfaces. Second, the \texttt{DataOut} class that is responsible
  for converting nodal data into information that can then be written
  into files for visualization did not know how to deal with vector-
  and tensor-valued fields whose components are complex numbers. An
  example for this is to solve the time-harmonic version of the
  Maxwell equations that has the electric and magnetic fields as
  solution. This, too, has been addressed in this release.
\item The class \texttt{DiscreteTime} is introduced to provide a more
  consistent, more readable, and less error-prone approach to control
  time-stepping algorithms within time-dependent simulations.
  The mutating interface of this class is designed to be minimal
  to enforce a number of important programming invariants, reducing
  the possibility of mistakes in the user code.
  When time-incrementation is requested within the user code, the class makes
  sure that time increases by a non-zero step size and the current step
  number is incremented accordingly.
  Furthermore, \texttt{DiscreteTime} ensures that the final time step ends
  precisely on a predefined end time. Therefore, the final step size is
  automatically lengthened or shortened to accommodate this feature.
  In addition, the class provides useful access methods which return the step
  number $n$, the values of the simulation time corresponding to $t_{n-1}$,
  $t_n$, and $t_{n+1}$, and the step-size values $t_n - t_{n-1}$ and
  $t_{n+1} - t_n$.
\item z \todo[inline]{What else? Maybe mention the updated step-12?}
\end{itemize}
%
In addition to these changes, the changelog lists more than 200 other
features and bugfixes.

@@

By replacing the collective communications during set up and removing the arrays
that contain information for each process (enabled by the application of consensus
algorithms and other modifications---a full list of modifications leading to this
improvement can be found online), we were able to significantly
improve the set up
time for large-scale simulations and to solve a Poisson problem with multigrid
with \num{2.1e12} unknowns.
Figure~\ref{fig:init_costs} compares the timings of simulations of various problem
sizes (including set up) on 49,152 MPI ranks using a matrix-free solver according
to~\cite{KronbichlerKormann2019,KronbichlerWall2018} discontinuous elements of
degree $5$ in a geometric multigrid (GMG) scheme with the previous release 9.1
and with the current release 9.2. While the scaling for the V-cycle had been
very good before, many initialization routines have been considerably
improved, especially the enumeration of unknowns on the multigrid levels and
the setup of the multigrid transfer.

@@

\begin{figure}
  \begin{tikzpicture}
    \begin{loglogaxis}[
      width=0.51\textwidth,
      height=0.38\textwidth,
          xlabel={\# DoFs},
          ylabel={time [s]},
          title={9.1 release},
          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
          tick label style={font=\scriptsize},
          xticklabel style={rotate=25,anchor=north east},
          label style={font=\scriptsize},
          legend style={font=\scriptsize},
          %legend to name=legend:scalingMG,
          %legend cell align=left,
          %legend columns = 3,
          ymin=2e-3,ymax=8e1,
          grid, thick]
          \addplot table[x index = 0, y index={1}] {\tableinitold};
          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinitold};
          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinitold};
          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinitold};
          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinitold};
          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinitold};
    \end{loglogaxis}
  \end{tikzpicture}
  \hfill
  \begin{tikzpicture}
    \begin{loglogaxis}[
      width=0.51\textwidth,
      height=0.38\textwidth,
          xlabel={\# DoFs},
          ylabel={time [s]},
          title={9.2 release},
          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
          tick label style={font=\scriptsize},
          xticklabel style={rotate=25,anchor=north east}, thick,
          label style={font=\scriptsize},
          legend style={font=\scriptsize},
          legend to name=legend:init,
          legend cell align=left,
          legend columns = 6,
          ymin=2e-3,ymax=8e1,
          grid]
          \addplot table[x index = 0, y index={1}] {\tableinit};
          \addlegendentry{create mesh};
          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinit};
          \addlegendentry{enumerate (mg) dofs};
          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinit};
          \addlegendentry{init matrix-free};
          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinit};
          \addlegendentry{init GMG transfer};
          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinit};
          \addlegendentry{init smoother};
          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinit};
          \addlegendentry{GMG V-cycle};
    \end{loglogaxis}
  \end{tikzpicture}
  \\
  \strut\hfill\pgfplotslegendfromname{legend:init}\hfill\strut
  \caption{Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right) when run on 49,152 MPI ranks.}
  \label{fig:init_costs}
\end{figure}

@@

Certain types of quantities in a simulation are constants fully known
at compile time. They can be pre-calculated and fully stored inside the
compiled program binary in order to avoid unnecessary initialization during
runtime. This optimization is now enabled for the class templates
\texttt{Tensor} and \texttt{SymmetricTensor} by qualifying their constructor,
member functions, and overloaded operators as \texttt{constexpr}.
For instance, the linear mechanical constitutive model for elastic solids
uses a constant fourth-order elasticity tensor
$\mathbb{C} = \lambda \boldsymbol{I} \otimes \boldsymbol{I} + 2 \mu \mathbb{I}$
which does not depend on the current state of strain.
This tensor can be statically initialized by defining it as
\texttt{constexpr SymmetricTensor<4, dim>}.
As another example, the lattice vectors in a crystal plasticity model are generally
constant and known during compilation time, enabling their efficient definition as
\texttt{constexpr Tensor<1, dim>}.
The capability of defining \texttt{constexpr} variables, functions, and methods
was introduced by the C++11 standard and was later expanded by the C++14 standard.
Therefore, the extent of \texttt{constexpr} support in \dealii{} depends on the C++
standard which is used to compile the library. The next release of \dealii{}
will fully adopt the features of the C++14 standard.

@@
